---
title: "PCA.EAS.str_length"
---

```{r knitr, include = FALSE}
DOCNAME <- "PCA.EAS.str_length"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

- First, run `PCA.EAS.str_length.R`

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# plot
library(scales)
library(patchwork)
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
library(ggrastr)
library(EnvStats)
theme_set(theme_cowplot(
  font_size = 12, # axis.title
  rel_small = 10 / 12, # axis.text
  rel_tiny = 9 / 12,
  rel_large = 12 / 12
))

# heatmap
library(pheatmap)

# color
library(ggsci)
library(RColorBrewer)

# data.table
library(data.table)
```

Load R source

```{r source, cache = FALSE}
source(here::here("code/color.R"))
```

Load sample info

```{r load-sam-info, cache=FALSE, message=FALSE, warning=FALSE}
sam_info <- read_tsv(here::here("data", "220116_sample_info.GangSTR.txt")) %>%
  dplyr::select(sample_old, dataset, super_pop, pop)
```

## Random 1 {.tabset}

Load PCA results

```{r load-pca-res, cache=FALSE, message=FALSE, warning=FALSE}
pc_res <- readRDS(here::here("output", DOCNAME, "random1.pc_res_df.rds"))
pc_var <- readRDS(here::here("output", DOCNAME, "random1.pc_var_df.rds"))
head(pc_res)
```

Neat

```{r}
d4p <- pc_res %>%
  left_join(sam_info, by = "sample_old")
```

Screen plot

```{r screen-1, fig.width=6, fig.height=4}
var_explained = pc_var %>%
  filter(pcs < 11) %>%
  mutate(Proportion = Proportion*100,
         pcs = factor(pcs))
var_explained

var_explained %>%
  ggplot(aes(x = pcs, y = Proportion, group = 1)) +
  geom_col(fill = "#4682b4") +
  geom_point_rast() +
  geom_line() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    y = "% of explained variances",
    x = "Dimensions"
  ) +
  theme_cowplot()
```

### By dataset {.tabset}

#### PC1 and PC2

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc1, y = pc2, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC2 (0.48%)")
```

#### PC1 and PC3

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc1, y = pc3, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC3 (0.43%)")
```

#### PC2 and PC3

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc2, y = pc3, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC2 (0.48%)", y = "PC3 (0.43%)")
```

#### PC3 and PC4

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc3, y = pc4, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC3 (0.43%)", y = "PC4 (0.32%)")
```

### By pop {.tabset}

#### PC1 and PC2

```{r, fig.height=6, fig.width=6}
p1.PC1PC2 = d4p %>%
  ggplot(aes(x = pc1, y = pc2, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC2 (0.48%)", color=NULL) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  theme(legend.position = "bottom")
p1.PC1PC2
```

#### PC1 and PC3

```{r, fig.height=6, fig.width=6}
p1.PC1PC3 = d4p %>%
  ggplot(aes(x = pc1, y = pc3, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC3 (0.43%)", color=NULL) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  theme(legend.position = "bottom")
p1.PC1PC3
```

#### PC2 and PC3

```{r, fig.height=6, fig.width=8}
d4p %>%
  ggplot(aes(x = pc2, y = pc3, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC2 (0.48%)", y = "PC3 (0.43%)")
```

#### PC3 and PC4

```{r, fig.height=6, fig.width=8}
d4p %>%
  ggplot(aes(x = pc3, y = pc4, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC3 (0.43%)", y = "PC4 (0.32%)")
```

## Random 2 {.tabset}

Load PCA results

```{r load-pca-res-2, cache=FALSE, message=FALSE, warning=FALSE}
pc_res <- readRDS(here::here("output", DOCNAME, "random2.pc_res_df.rds"))
pc_var <- readRDS(here::here("output", DOCNAME, "random2.pc_var_df.rds"))
head(pc_res)
```

Neat

```{r}
d4p <- pc_res %>%
  left_join(sam_info, by = "sample_old")
```

Screen plot

```{r screen-2, fig.width=6, fig.height=4}
var_explained = pc_var %>%
  filter(pcs < 11) %>%
  mutate(Proportion = Proportion*100,
         pcs = factor(pcs))
var_explained

var_explained %>%
  ggplot(aes(x = pcs, y = Proportion, group = 1)) +
  geom_col(fill = "#4682b4") +
  geom_point_rast() +
  geom_line() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    y = "% of explained variances",
    x = "Dimensions"
  ) +
  theme_cowplot()
```

### By dataset

#### PC1 and PC2 {.tabset}

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc1, y = pc2, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC2 (0.48%)")
```

#### PC1 and PC3

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc1, y = pc3, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC3 (0.43%)")
```

#### PC2 and PC3

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc2, y = pc3, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC2 (0.48%)", y = "PC3 (0.43%)")
```

#### PC3 and PC4

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc3, y = pc4, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC3 (0.43%)", y = "PC4 (0.31%)")
```

### By pop {.tabset}

#### PC1 and PC2

```{r, fig.height=6, fig.width=6}
p2.PC1PC2 = d4p %>%
  ggplot(aes(x = pc1, y = pc2, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC2 (0.48%)", color=NULL) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  theme(legend.position = "bottom")
p2.PC1PC2
```

#### PC1 and PC3

```{r, fig.height=6, fig.width=6}
p2.PC1PC3 = d4p %>%
  ggplot(aes(x = pc1, y = pc3, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC3 (0.43%)", color=NULL) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  theme(legend.position = "bottom")
p2.PC1PC3
```

#### PC2 and PC3

```{r, fig.height=6, fig.width=8}
d4p %>%
  ggplot(aes(x = pc2, y = pc3, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC2 (0.48%)", y = "PC3 (0.43%)")
```

#### PC3 and PC4

```{r, fig.height=6, fig.width=8}
d4p %>%
  ggplot(aes(x = pc3, y = pc4, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC3 (0.43%)", y = "PC4 (0.31%)")
```

## Random 3 {.tabset}

Load PCA results

```{r load-pca-res-3, cache=FALSE, message=FALSE, warning=FALSE}
pc_res <- readRDS(here::here("output", DOCNAME, "random3.pc_res_df.rds"))
pc_var <- readRDS(here::here("output", DOCNAME, "random3.pc_var_df.rds"))
head(pc_res)
```

Neat

```{r}
d4p <- pc_res %>%
  left_join(sam_info, by = "sample_old")
```

Screen plot

```{r screen-3, fig.width=6, fig.height=4}
var_explained = pc_var %>%
  filter(pcs < 11) %>%
  mutate(Proportion = Proportion*100,
         pcs = factor(pcs))
var_explained

var_explained %>%
  ggplot(aes(x = pcs, y = Proportion, group = 1)) +
  geom_col(fill = "#4682b4") +
  geom_point_rast() +
  geom_line() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    y = "% of explained variances",
    x = "Dimensions"
  ) +
  theme_cowplot()
```

### By dataset {.tabset}

#### PC1 and PC2

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc1, y = pc2, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC2 (0.47%)")
```

#### PC1 and PC3

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc1, y = pc3, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC3 (0.43%)")
```

#### PC2 and PC3

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc2, y = pc3, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC2 (0.47%)", y = "PC3 (0.43%)")
```

#### PC3 and PC4

```{r, fig.height=6, fig.width=7}
d4p %>%
  ggplot(aes(x = pc3, y = pc4, color = dataset)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC3 (0.43%)", y = "PC4 (0.31%)")
```

### By pop {.tabset}

#### PC1 and PC2

```{r, fig.height=6, fig.width=6}
p3.PC1PC2 = d4p %>%
  ggplot(aes(x = pc1, y = pc2, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC2 (0.47%)", color=NULL) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  theme(legend.position = "bottom")
p3.PC1PC2
```

#### PC1 and PC3

```{r, fig.height=6, fig.width=6}
p3.PC1PC3 = d4p %>%
  ggplot(aes(x = pc1, y = pc3, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC1 (0.85%)", y = "PC3 (0.43%)", color=NULL) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  theme(legend.position = "bottom")
p3.PC1PC3
```

#### PC2 and PC3

```{r, fig.height=6, fig.width=8}
d4p %>%
  ggplot(aes(x = pc2, y = pc3, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC2 (0.47%)", y = "PC3 (0.43%)")
```

#### PC3 and PC4

```{r, fig.height=6, fig.width=8}
d4p %>%
  ggplot(aes(x = pc3, y = pc4, color = pop)) +
  geom_point_rast() +
  scale_color_manual(values = pop_color_maps) +
  labs(x = "PC3 (0.43%)", y = "PC4 (0.31%)")
```

## Figures

Random1

```{r fig.width=7.5, fig.height=4.5}
p.PCA.EAS.random1 = p1.PC1PC2 + p1.PC1PC3 +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
p.PCA.EAS.random1
```

Random2

```{r fig.width=7.5, fig.height=4.5}
p.PCA.EAS.random2 = p2.PC1PC2 + p2.PC1PC3 +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
p.PCA.EAS.random2
```

Random3

```{r fig.width=7.5, fig.height=4.5}
p.PCA.EAS.random3 = p3.PC1PC2 + p3.PC1PC3 +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
p.PCA.EAS.random3
```

## Save

```{r save, cache=FALSE}
saveRDS(p.PCA.EAS.random1, file = here::here("output", DOCNAME, "Fig.PCA.EAS.random1.rds"))
saveRDS(p.PCA.EAS.random2, file = here::here("output", DOCNAME, "Fig.PCA.EAS.random2.rds"))
saveRDS(p.PCA.EAS.random3, file = here::here("output", DOCNAME, "Fig.PCA.EAS.random3.rds"))
```

## Session info


