---
title: "sample_callrate"
---

```{r knitr, include = FALSE}
DOCNAME = "sample_callrate"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# ggplot
library(scales)
library(ggside)
library(ggpmisc)
library(patchwork)
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
library(ggrastr)
theme_set(theme_cowplot(
  font_size = 10, # Overall font size
  rel_small = 10 / 10, # axis tick labels
  rel_tiny = 8 / 10, # caption
  rel_large = 10 / 10 # title
))

# heatmap
library(pheatmap)

# color
library(ggsci)
library(RColorBrewer)

# data.table
library(data.table)
```

Load R source

```{r source, cache = FALSE}
source(here::here("code/color.R"))
```

Load sample info

```{r load-sam-info, cache=FALSE, message=FALSE, warning=FALSE}
sam_info = read_tsv(here::here('data', '220116_sample_info.GangSTR.txt'))
```

Load callrate

```{r load-sam-callrate, cache=FALSE, warning=FALSE, message=FALSE}
sam_callrate = read_tsv('/niuyw-usb-disk/Projects/STR/str_stat/sample_callrate.txt')
```

Load new sample coverage (by mosdepth)

```{r load-sam-cov, cache=FALSE, warning=FALSE, message=FALSE}
sam_cov = read_csv(here::here('data', 'sam8772.mosdepth.csv')) %>%
  dplyr::select(sample_old = sam_old, coverage2 = depth.autosome)
```

## Neat

```{r neat}
d4p = sam_info %>%
  left_join(sam_cov, by = "sample_old") %>%
  left_join(sam_callrate, by = "sample_old") %>%
  dplyr::select(-coverage) %>%
  mutate(callrate = called/765227*100)
```

## Stat

Cov

```{r}
# mean
mean(d4p$coverage2)
median(d4p$coverage2)
sd(d4p$coverage2)
# range
range(d4p$coverage2)

# mean by dataset
d4p %>%
  group_by(dataset) %>%
  summarise(mean_cov = mean(coverage2),
            median_cov = median(coverage2),
            sd_cov = sd(coverage2))

# range by dataset
d4p %>%
  group_by(dataset) %>%
  summarise(r = range(coverage2))
```

The call rates of over 95% samples were larger than 95%.

```{r}
d4p %>%
  mutate(lt90 = ifelse(callrate > 90, "Yes", "No")) %>%
  group_by(lt90) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n))
```

Call rate

```{r}
mean(d4p$callrate)

d4p %>%
  group_by(dataset) %>%
  summarise(m = mean(callrate))
```

Num called

```{r}
mean(d4p$called)

d4p %>%
  group_by(dataset) %>%
  summarise(m = mean(called))
```

## Sample coverage {.tabset}

### Histgram

All samples

```{r fig.width=3.5, fig.height=3}
d4p %>%
  ggplot(aes(x = coverage2)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 31.9, linetype = "dashed") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Autosome coverage", y ="# of samples")
```

NyuWa

```{r fig.width=3.5, fig.height=3}
p.cov.NyuWa = d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = coverage2)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 31.01	, linetype = "dashed") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Autosome coverage", y ="# of samples")
p.cov.NyuWa

# save
ggsave(p.cov.NyuWa, filename = here::here("output", DOCNAME, "Fig.cov.NyuWa.pdf"),
       height = 3, width = 3.5)
saveRDS(p.cov.NyuWa, file = here::here("output", DOCNAME, "Fig.cov.NyuWa.rds"))
```

1KGP

```{r fig.width=3.5, fig.height=3}
p.cov.1KGP = d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = coverage2)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 32.54	, linetype = "dashed") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Autosome coverage", y ="# of samples")
p.cov.1KGP

# save
ggsave(p.cov.1KGP, filename = here::here("output", DOCNAME, "Fig.cov.1KGP.pdf"),
       height = 3, width = 3.5)
saveRDS(p.cov.1KGP, file = here::here("output", DOCNAME, "Fig.cov.1KGP.rds"))
```

### Density

All samples

```{r fig.width=3.5, fig.height=3}
d4p %>%
  ggplot(aes(x = coverage2, fill = dataset)) +
  geom_density(alpha = .5) +
  geom_vline(xintercept = c(31.01, 32.54), linetype = "dashed") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Autosome coverage") +
  scale_fill_manual(values = c("#7da1d5", "#bdde85")) +
  theme(legend.position = c(0.7, 0.2))
```

NyuWa

```{r fig.width=3.5, fig.height=3}
p = d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = coverage2)) +
  geom_density(fill= "lightblue") +
  geom_vline(xintercept = 31.01	, linetype = "dashed") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Autosome coverage")
p
```

1KGP

```{r fig.width=3.5, fig.height=3}
p = d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = coverage2)) +
  geom_density(fill= "lightblue") +
  geom_vline(xintercept = 32.54	, linetype = "dashed") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Autosome coverage")
p
```

## Num called {.tabset}

### Num called and coverage

All

```{r fig.width=3.5, fig.height=3}
d4p %>%
  ggplot(aes(x = coverage2, y = called, color = dataset)) +
  geom_point_rast(size = 0.7, raster.dpi = 600) +
  scale_color_manual(values = c("#7da1d5", "#bdde85")) +
  scale_y_continuous(labels = label_comma()) +
  labs(y = "# of called", x= "Autosome coverage") +
  theme(legend.position = c(0.7, 0.2))
```

NyuWa

```{r fig.width=3.5, fig.height=3}
d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = coverage2, y = called)) +
  geom_point_rast(size = 0.7, raster.dpi = 600) +
  scale_y_continuous(labels = label_comma()) +
  labs(y = "# of called", x= "Autosome coverage")
```

1KGP

```{r fig.width=3.5, fig.height=3}
d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = coverage2, y = called)) +
  geom_point_rast(size = 0.7, raster.dpi = 600) +
  scale_y_continuous(labels = label_comma()) +
  labs(y = "# of called", x= "Autosome coverage")
```

### Density

All

```{r fig.width=3, fig.height=2.5}
p = d4p %>%
  ggplot(aes(x = called)) +
  geom_density(fill= "lightblue") +
  geom_rug(aes(x = called, y = 0), position = position_jitter(height = 0)) +
  scale_x_continuous(expand = c(0, 0),
                     labels = label_comma()) +
  labs(x = "# of called")
p
```

NyuWa

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = called)) +
  geom_density(fill= "lightblue") +
  geom_rug(aes(x = called, y = 0), position = position_jitter(height = 0)) +
  scale_x_continuous(expand = c(0, 0),
                     labels = label_comma()) +
  labs(x = "# of called")
```

1KGP

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = called)) +
  geom_density(fill= "lightblue") +
  geom_rug(aes(x = called, y = 0), position = position_jitter(height = 0)) +
  scale_x_continuous(expand = c(0, 0),
                     labels = label_comma()) +
  labs(x = "# of called")
```

### Histgram

All

```{r  fig.width=4, fig.height=3.5}
d4p %>%
  ggplot(aes(x = called)) +
  geom_histogram() +
  #scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "# of called", y ="# of samples")
```

NyuWa

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = called)) +
  geom_histogram() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "# of called", y ="# of samples")
```

1KGP

```{r fig.width=3.5, fig.height=3}
d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = called)) +
  geom_histogram() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "# of called", y ="# of samples")
```

## Callrate {.tabset}

### Callrate and coverage

All

```{r fig.width=3.5, fig.height=3}
d4p %>%
  ggplot(aes(x = coverage2, y = callrate, color = dataset)) +
  geom_point_rast(size = 0.7, raster.dpi = 600) +
  scale_color_manual(values = c("#7da1d5", "#bdde85")) +
  labs(y = "Sample call rate (%)", x= "Autosome coverage") +
  theme(legend.position = c(0.7, 0.2))
```

NyuWa

```{r fig.width=3.5, fig.height=3}
p.callrate_cov.NyuWa = d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = coverage2, y = callrate)) +
  geom_point_rast(size = 0.5, raster.dpi = 600, color = "#1f77b4") +
  geom_xsideboxplot(aes(y = dataset), orientation = "y",
                    outlier.size = 0.02, color = "#1f77b4") +
  scale_xsidey_discrete(labels = NULL) +
  geom_ysideboxplot(aes(x = dataset), orientation = "x",
                    outlier.size = 0.02, color = "#1f77b4") +
  scale_ysidex_discrete(guide = guide_axis(angle = 45),
                        labels = NULL) +
  stat_correlation(label.y = "bottom",
                   label.x = "right") +
  labs(y = "Sample call rate (%)", x= "Autosome coverage")
p.callrate_cov.NyuWa
```

1KGP

```{r fig.width=3.5, fig.height=3}
p.callrate_cov.1KGP = d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = coverage2, y = callrate)) +
  geom_point_rast(size = 0.5, raster.dpi = 600, color = "#1f77b4") +
  geom_xsideboxplot(aes(y = dataset), orientation = "y",
                    outlier.size = 0.02, color = "#1f77b4") +
  scale_xsidey_discrete(labels = NULL) +
  geom_ysideboxplot(aes(x = dataset), orientation = "x",
                    outlier.size = 0.02, color = "#1f77b4") +
  scale_ysidex_discrete(guide = guide_axis(angle = 45),
                        labels = NULL) +
  stat_correlation(label.y = "bottom",
                   label.x = "right") +
  labs(y = "Sample call rate (%)", x= "Autosome coverage")
p.callrate_cov.1KGP
```

### Density

All

```{r fig.width=3, fig.height=2.5}
p = d4p %>%
  ggplot(aes(x = callrate)) +
  geom_density(fill= "lightblue") +
  geom_rug(aes(x = callrate, y = 0), position = position_jitter(height = 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(x = "Call rate (%)")
p
ggsave(p, filename = here::here("output", DOCNAME, "sample_callrate.pdf"),
       height = 3.5, width = 4)
```

NyuWa

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = callrate)) +
  geom_density(fill= "lightblue") +
  geom_rug(aes(x = callrate, y = 0), position = position_jitter(height = 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(x = "Call rate (%)")
```

1KGP

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = callrate)) +
  geom_density(fill= "lightblue") +
  geom_rug(aes(x = callrate, y = 0), position = position_jitter(height = 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(x = "Call rate (%)")
```

### Histgram

All

```{r  fig.width=4, fig.height=3.5}
d4p %>%
  ggplot(aes(x = callrate)) +
  geom_histogram(binwidth=.05) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Call rate (%)", y ="# of samples")
```

NyuWa

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == 'NyuWa') %>%
  ggplot(aes(x = callrate)) +
  geom_histogram(binwidth=.05) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Call rate (%)", y ="# of samples")
```

1KGP

```{r fig.width=4, fig.height=3.5}
d4p %>%
  filter(dataset == '1KGP') %>%
  ggplot(aes(x = callrate)) +
  geom_histogram(binwidth=.05, colour="black") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Call rate (%)", y ="# of samples")
```

## Align plots

### Coverage

```{r fig.height=3, fig.width=6}
p.cov = p.cov.NyuWa + p.cov.1KGP +
  plot_layout(ncol = 2)
p.cov

# save
ggsave(p.cov, filename = here::here("output", DOCNAME, "Fig.cov.NyuWa_1KGP.pdf"),
       height = 3, width = 6)
saveRDS(p.cov, file = here::here("output", DOCNAME, "Fig.cov.NyuWa_1KGP.rds"))
```

### Callrate and coverage

```{r fig.height=3, fig.width=6}
p.callrate_cov = p.callrate_cov.NyuWa + p.callrate_cov.1KGP +
  plot_layout(ncol = 2)
p.callrate_cov

# save
ggsave(p.callrate_cov, filename = here::here("output", DOCNAME, "Fig.callrate_cov.NyuWa_1KGP.pdf"),
       height = 3, width = 6)
saveRDS(p.callrate_cov, file = here::here("output", DOCNAME, "Fig.callrate_cov.NyuWa_1KGP.rds"))
```

## Save

```{r save, cache=FALSE}
saveRDS(d4p, file = here::here("output", DOCNAME, "sample_callrate.rds"))
```

## Session info


