---
title: "aSTR.preprocess_pudi"
---

```{r knitr, include = FALSE}
DOCNAME <- "aSTR.preprocess_pudi"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

- 处理 Geuvadis 项目的 DaPars2 数据

- 用于 QTL 分析

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# plot
library(scales)
library(patchwork)
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
library(ggrastr)
library(EnvStats)
theme_set(theme_cowplot(
  font_size = 12, # axis.title
  rel_small = 10 / 12, # axis.text
  rel_tiny = 9 / 12,
  rel_large = 12 / 12
))

# heatmap
library(pheatmap)
library(corrplot)

# color
library(ggsci)
library(RColorBrewer)

# data.table
library(data.table)

# hidden factor
library(SummarizedExperiment)
library(iasva)
```

Load R source

```{r source, cache = FALSE}
source(here::here("code/color.R"))
```

Load PCA eigenvec

```{r load-eigenvec, cache=FALSE, warning=FALSE, message=FALSE}
pop_pc10 <- read_delim("/niuyw-usb-disk/Projects/Geuvadis/geno_pca/autosome_maf0.05_pca.eigenvec",
  delim = " ",
  col_names = FALSE
) %>%
  dplyr::select(-X2)
```

Load sample info

```{r load-sam-info, message=FALSE, warning=FALSE, cache=FALSE}
sam_info <- read_tsv("/niuyw-usb-disk/Projects/Geuvadis/count/targets.txt") %>%
  column_to_rownames("sample")
```

Load RAN-seq exp

```{r load-exp, cache=FALSE, warning=FALSE, message=FALSE}
# load exp
rna_exp <- readRDS(here::here("output", "eSTR.preprocess_exp", "Geuvadis.exp_adjusted.rds"))
rna_exp[1:3, 1:3]
```

Load DaPars2 result

```{r load-pudi, cache=FALSE, warning=FALSE, message=FALSE}
df.DaPars2 = read_delim("/niuyw-usb-disk/Projects/Geuvadis/DaPars2/DaPars2.res.txt", delim = "\t")
df.DaPars2[1:3, 1:5]
```

## Neat

Filter samples

```{r filter-samples}
sam_info <- sam_info[pop_pc10$X1, ]
table(sam_info$sex)
table(sam_info$pop)
table(sam_info$spop)
```

Get PUDI table, and filter

```{r neat-pudi}
df.pudi = df.DaPars2 %>%
  column_to_rownames("Gene")
# 445 samples
df.pudi = df.pudi[, rownames(rna_exp)]
dim(df.pudi)

# transcripts with missing values in > 50% of individuals were removed.
df.pudi = df.pudi[rowSums(is.na(df.pudi)) < 445/2, ]
dim(df.pudi)

# remove rows with SD = 0
x = apply(df.pudi, 1, function(i){sd(i, na.rm = TRUE)})
df.pudi = df.pudi[x != 0,]
dim(df.pudi)
```

Test if we could keep genes used eSTR analysis. The number is OK.

```{r test-filter-by-gene}
df.pudi %>%
  rownames_to_column("Gene") %>%
  mutate(gene = str_split(Gene, "[|]", simplify = TRUE)[,2]) %>%
  filter(gene %in% colnames(rna_exp)) %>%
  dim()
```

Keep only genes used in eSTR analysis

```{r filter-pudi-by-gene}
df.pudi = df.pudi %>%
  rownames_to_column("Gene") %>%
  mutate(gene = str_split(Gene, "[|]", simplify = TRUE)[,2]) %>%
  filter(gene %in% colnames(rna_exp)) %>%
  dplyr::select(-gene) %>%
  column_to_rownames("Gene")
dim(df.pudi)

# use transcript as rowname
rownames(df.pudi) = str_split(rownames(df.pudi), "[|]", simplify = TRUE)[,1]
df.pudi[1:3, 1:3]
```

## Run IA-SVA

With the following covariates

- sex

- 10 PCs for population structure

```{r run-iasva}
set.seed(100)

# known factors
mod <- model.matrix(~ sam_info$sex +
  pop_pc10$X3 + pop_pc10$X4 + pop_pc10$X5 + pop_pc10$X6 + pop_pc10$X7 +
  pop_pc10$X8 + pop_pc10$X9 + pop_pc10$X10 + pop_pc10$X11 + pop_pc10$X12)

# create a summarizedexperiment class
summ_exp <- SummarizedExperiment(assays = na.omit(df.pudi))

# run
iasva.res <- iasva(summ_exp, mod[, -1],
  verbose = FALSE,
  permute = FALSE, num.sv = 20
)
```

## Plot

Get res

```{r get-res}
iasva.sv <- iasva.res$sv
iasva.sv[1:3, 1:3]
```

Correlation of SVs

```{r , fig.width=4, fig.height=5}
corrplot(cor(iasva.sv), type = "lower",
         method = "square", tl.cex = 0.8)
pdf(here::here("output", DOCNAME, "Fig.iasva_sv.corrplot.pdf"),
    height = 5, width = 4)
corrplot(cor(iasva.sv), type = "lower", method = "square", tl.cex = 0.8)
dev.off()
```

Plot SVs

```{r, fig.width=3.5, fig.height=5}
d4p = cbind(sam_info, as.data.frame(iasva.sv))

p.SV1SV2 = d4p %>%
  ggplot(aes(x = SV1, y = SV2, color = pop)) +
  geom_point_rast() +
  scale_color_jco()
p.SV3SV4 = d4p %>%
  ggplot(aes(x = SV3, y = SV4, color = pop)) +
  geom_point_rast() +
  scale_color_jco()

p.SV.sample_scatter = p.SV1SV2 + p.SV3SV4 +
  plot_layout(guides = "collect", ncol = 1)
p.SV.sample_scatter

ggsave(p.SV.sample_scatter, filename = here::here("output", DOCNAME, "Fig.iasva_sv.sample_scatter.pdf"),
       height = 5, width = 3.5)
saveRDS(p.SV.sample_scatter, file = here::here("output", DOCNAME, "Fig.iasva_sv.sample_scatter.rds"))
```

## Adjust pudi

Prepare dat for linear model

```{r prepare-dat}
# exp
dat <- as.data.frame(t(df.pudi))
# PCA
dat <- cbind(pop_pc10[, -1], dat)
colnames(dat)[1:10] <- paste0("PC", 1:10)
# sex
dat$sex <- sam_info$sex
# SV
dat <- cbind(dat, as.data.frame(iasva.sv[, 1:15]))
```

Linear model to adjust

```{r linear-model, eval=FALSE}
# genes
genes <- rownames(df.pudi)

# formula
univ_formulas <- sapply(genes, function(y) {
  #print(y)
  as.formula(paste0(y, " ~ sex+PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+SV1+SV2+SV3+SV4+SV5+SV6+SV7+SV8+SV9+SV10+SV11+SV12+SV13+SV14+SV15"))
})

# adjust exp
dat_adj <- sapply(univ_formulas, function(f) {
  # print(f)
  lm.res <- lm(f, data = dat, na.action = na.exclude)
  residuals(lm.res)
})

# to data.frame
dat_adj = as.data.frame(dat_adj)
dat_adj[1:3, 1:3]
```

## Save

```{r save, cache=FALSE, eval=FALSE}
saveRDS(df.pudi, file = here::here("output", DOCNAME, "Geuvadis.df.pudi.rds"))
saveRDS(iasva.res, file = here::here("output", DOCNAME, "Geuvadis.iasva.res.rds"))
saveRDS(dat_adj, file = here::here("output", DOCNAME, "Geuvadis.pudi_adjusted.rds"))
```


