---
title: "eSTR.identification"
---

```{r knitr, include = FALSE}
DOCNAME <- "eSTR.identification"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

- First, run `eSTR.identification.R` 和 `eSTR.identification.permuted.R`

- Neat eSTR identification results

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# plot
library(scales)
library(patchwork)
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
library(ggrastr)
library(EnvStats)
theme_set(theme_cowplot(
  font_size = 10, # Overall font size
  rel_small = 10 / 10, # axis tick labels
  rel_tiny = 8 / 10, # caption
  rel_large = 10 / 10 # title
))

# heatmap
library(pheatmap)
library(corrplot)

# color
library(ggsci)
library(RColorBrewer)
```

Load R source

```{r source, cache = FALSE}
source(here::here("code/color.R"))
```

## Load data

Load exp

```{r load-exp, cache=FALSE, warning=FALSE, message=FALSE}
rna_exp = readRDS(here::here("output", "eSTR.preprocess_exp", "Geuvadis.exp_adjusted.rds"))
rna_exp[1:3, 1:3]
```

Load STR geno

```{r load-STR-geno, cache=FALSE, warning=FALSE, message=FALSE}
str_geno = read_csv("/niuyw-usb-disk/Projects/STR/eSTR/Geuvadis_445.str_geno.csv") %>%
  column_to_rownames("site")
str_geno[1:3, 1:3]
```

Load STR-gene pairs

```{r load-STR-gene-pairs, message=FALSE, warning=FALSE, cache=FALSE}
df.str_gene_pair = read_csv("/niuyw-usb-disk/Projects/STR/eSTR/Geuvadis_445.str_gene_pair.csv")
head(df.str_gene_pair)
```

Load eSTR result

```{r load-eSTR-res, message=FALSE, warning=FALSE, cache=FALSE}
fit_res = readRDS(here::here("output", DOCNAME, "eSTR.exp_str.glm_fit.rds"))
fit_res.permuted = readRDS(here::here("output", DOCNAME, "eSTR.exp_str.glm_fit.permuted.rds"))
```

## Neat

No. of pairs

```{r pairs}
dim(df.str_gene_pair)
```

No. of genes

```{r gene-str-neat}
# gene-str lst
gene_str_lst = split(df.str_gene_pair$site, df.str_gene_pair$gene_id)
length(gene_str_lst)
mean(sapply(gene_str_lst, length, simplify = TRUE))
```

No. of STRs

```{r STR-gene-neat}
# gene-str lst
str_gene_lst = split(df.str_gene_pair$gene_id, df.str_gene_pair$site)
length(str_gene_lst)
mean(sapply(str_gene_lst, length, simplify = TRUE))
```

eSTRs table.

> We used a gene-level FDR threshold (described previously14) of 10% to identify significant STR–gene pairs. We assume most genes have, at most, a single causal eSTR. For each gene, we determined the STR association with the strongest P value. This P value was adjusted using a Bonferroni correction for the number of STRs tested per gene, to give a P value for observing a single eSTR association for each gene. We then used the list of adjusted P values (one per gene) as input to the fdrcorrection function in the statsmodels.stats.multitest module to obtain a q-value for the best eSTR for each gene. [(Fotsing et al., 2019)](https://www.nature.com/articles/s41588-019-0521-9)

```{r eSTR-neat}
# get ONE str for each gene
tmp.lst = lapply(fit_res, function(x){
  x[1,]
})

df.eSTR = do.call(rbind, tmp.lst) %>%
  mutate(final.p = p.adjust(q.value, method = "BH")) %>%
  arrange(final.p)
head(df.eSTR)
```

## Stat

FDR 0.05

```{r stat-fdr0.05, fig.width=5, fig.height=5}
df.eSTR %>%
  filter(final.p < 0.05) %>%
  dim()
```

FDR 0.1

```{r stat-fdr0.1, fig.width=5, fig.height=5}
# total pairs
df.eSTR %>%
  filter(final.p < 0.1) %>%
  dim()

# genes
df.eSTR %>%
  filter(final.p < 0.1) %>%
  pull(gene) %>%
  unique() %>%
  length()

# STRs
df.eSTR %>%
  filter(final.p < 0.1) %>%
  pull(site) %>%
  unique() %>%
  length()
```

## Plot

### Q-Q plot

```{r qqplot, fig.width=2.5, fig.height=2.5}
df.fit_res.permuted = do.call(rbind, fit_res.permuted)
df.fit_res = do.call(rbind, fit_res)

# Observed data
y = sort(-log10(df.fit_res$p.value))
x = sort(-log10(runif(length(y))))

# Control data
yctrl = sort(-log10(df.fit_res.permuted$p.value))
xctrl = sort(-log10(runif(length(yctrl))))

# plot
d4p = data.frame(y = y,
                 x = x,
                 yctrl = yctrl,
                 xctrl = xctrl)
p = d4p %>%
  ggplot(aes(x=x, y=y)) +
  geom_point_rast(color="red", size = .7, raster.dpi = 600, dev = "ragg") +
  geom_point_rast(aes(x = xctrl, y = yctrl),
                  color = "gray", size = .7, raster.dpi = 600, dev = "ragg") +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Expected P value (-log10)",
       y = "Observed P value (-log10)")
p
ggsave(p, filename = here::here("output", DOCNAME, "Fig.qqplot.pdf"),
       height = 2.5, width = 2.5)
saveRDS(p, file = here::here("output", DOCNAME, "Fig.qqplot.rds"))
```

### eSTR examples {.tabset}

#### chr3:63912686

SCA7

> chr3	63912686	63912715	CAG	SCA7

```{r fig.width=4, fig.height=4}
s = "chr3:63912686"
g = "ENSG00000285399.3"
s_geno = na.omit(unlist(str_geno[s, ]))
g_exp = rna_exp[, g]
names(g_exp) = rownames(rna_exp)
g_exp = g_exp[names(s_geno)]
dat = data.frame(sample = names(g_exp), exp = g_exp, geno = factor(s_geno))

dat %>%
  ggplot(aes(x = geno, y = exp)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(width = .4, outlier.colour = "black", outlier.size = .5) +
  stat_summary(fun=mean, geom="line", aes(group=1), colour="red", size=1) + 
  stat_summary(fun=mean, geom="point", colour="red", size=1.5, shape = 15) +
  geom_jitter_rast(position=position_jitter(0.2), size = .5,
              dev = "ragg", raster.dpi = 600) +
  stat_n_text() +
  labs(x = "STR dosage", y = "Normalized expression",
       title = paste(g, s, sep = " - "))
```

#### chr10:45384834 - ALOX5

```{r fig.width=4, fig.height=4}
s = "chr10:45384834"
g = "ENSG00000012779.11"
s_geno = na.omit(unlist(str_geno[s, ]))
g_exp = rna_exp[, g]
names(g_exp) = rownames(rna_exp)
g_exp = g_exp[names(s_geno)]
dat = data.frame(sample = names(g_exp), exp = g_exp, geno = factor(s_geno))

dat %>%
  ggplot(aes(x = geno, y = exp)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(width = .4, outlier.colour = "black", outlier.size = .5) +
  stat_summary(fun=mean, geom="line", aes(group=1), colour="red", size=1) + 
  stat_summary(fun=mean, geom="point", colour="red", size=1.5, shape = 15) +
  geom_jitter_rast(position=position_jitter(0.2), size = .5,
              dev = "ragg", raster.dpi = 600) +
  stat_n_text() +
  labs(x = "STR dosage", y = "Normalized expression",
       title = paste(g, s, sep = " - "))
```

#### chr7:100359506

```{r fig.width=4, fig.height=4}
s = "chr7:100359506"
g = "ENSG00000121716.20"
s_geno = na.omit(unlist(str_geno[s, ]))
g_exp = rna_exp[, g]
names(g_exp) = rownames(rna_exp)
g_exp = g_exp[names(s_geno)]
dat = data.frame(sample = names(g_exp), exp = g_exp, geno = factor(s_geno))

dat %>%
  ggplot(aes(x = geno, y = exp)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(width = .4, outlier.colour = "black", outlier.size = .5) +
  stat_summary(fun=mean, geom="line", aes(group=1), colour="red", size=1) + 
  stat_summary(fun=mean, geom="point", colour="red", size=1.5, shape = 15) +
  geom_jitter_rast(position=position_jitter(0.2), size = .5,
              dev = "ragg", raster.dpi = 600) +
  stat_n_text() +
  labs(x = "STR dosage", y = "Normalized expression",
       title = paste(g, s, sep = " - "))
```

#### chr5:96936392

```{r fig.width=4, fig.height=4}
s = "chr5:96936392"
g = "ENSG00000164308.17"
s_geno = na.omit(unlist(str_geno[s, ]))
g_exp = rna_exp[, g]
names(g_exp) = rownames(rna_exp)
g_exp = g_exp[names(s_geno)]
dat = data.frame(sample = names(g_exp), exp = g_exp, geno = factor(s_geno))

dat %>%
  ggplot(aes(x = geno, y = exp)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(width = .4, outlier.colour = "black", outlier.size = .5) +
  stat_summary(fun=mean, geom="line", aes(group=1), colour="red", size=1) + 
  stat_summary(fun=mean, geom="point", colour="red", size=1.5, shape = 15) +
  geom_jitter_rast(position=position_jitter(0.2), size = .5,
              dev = "ragg", raster.dpi = 600) +
  stat_n_text() +
  labs(x = "STR dosage", y = "Normalized expression",
       title = paste(g, s, sep = " - "))
```

#### chr16:56472926

```{r fig.width=4, fig.height=4}
s = "chr16:56472926"
g = "ENSG00000125124.12"
s_geno = na.omit(unlist(str_geno[s, ]))
g_exp = rna_exp[, g]
names(g_exp) = rownames(rna_exp)
g_exp = g_exp[names(s_geno)]
dat = data.frame(sample = names(g_exp), exp = g_exp, geno = factor(s_geno))

dat %>%
  ggplot(aes(x = geno, y = exp)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(width = .4, outlier.colour = "black", outlier.size = .5) +
  stat_summary(fun=mean, geom="line", aes(group=1), colour="red", size=1) + 
  stat_summary(fun=mean, geom="point", colour="red", size=1.5, shape = 15) +
  geom_jitter_rast(position=position_jitter(0.2), size = .5,
              dev = "ragg", raster.dpi = 600) +
  stat_n_text() +
  labs(x = "STR dosage", y = "Normalized expression",
       title = paste(g, s, sep = " - "))
```

#### chr11:43772024

```{r fig.width=4, fig.height=4}
s = "chr11:43772024"
g = "ENSG00000149084.13"
s_geno = na.omit(unlist(str_geno[s, ]))
g_exp = rna_exp[, g]
names(g_exp) = rownames(rna_exp)
g_exp = g_exp[names(s_geno)]
dat = data.frame(sample = names(g_exp), exp = g_exp, geno = factor(s_geno))

dat %>%
  ggplot(aes(x = geno, y = exp)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(width = .4, outlier.colour = "black", outlier.size = .5) +
  stat_summary(fun=mean, geom="line", aes(group=1), colour="red", size=1) + 
  stat_summary(fun=mean, geom="point", colour="red", size=1.5, shape = 15) +
  geom_jitter_rast(position=position_jitter(0.2), size = .5,
              dev = "ragg", raster.dpi = 600) +
  stat_n_text() +
  labs(x = "STR dosage", y = "Normalized expression",
       title = paste(g, s, sep = " - "))
```

## Save

```{r save, cache=FALSE, eval=FALSE}
saveRDS(df.eSTR, file = here::here("output", DOCNAME, "df.eSTR.rds"))
write_csv(df.eSTR, file = here::here("output", DOCNAME, "df.eSTR.csv"))
```

