---
title: "str_cmp_NyuWa_1KGP"
---

```{r knitr, include = FALSE}
DOCNAME = "str_cmp_NyuWa_1KGP"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = FALSE,
    cache.lazy     = FALSE, 
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# plot
library(scales)
library(patchwork)
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
library(ggrastr)
library(EnvStats)
library(venneuler)

theme_set(theme_cowplot(
  font_size = 10, # Overall font size
  rel_small = 10 / 10, # axis tick labels
  rel_tiny = 8 / 10, # caption
  rel_large = 10 / 10 # title
))
geom.text.size <- 7

# heatmap
library(pheatmap)

# color
library(ggsci)
library(RColorBrewer)

# data.table
library(data.table)

## 2, 3, 4, 5, 6
cPalette <- c("#999999", "#7DA1D4", "#BEDE85", "#DCB070", "#EAAAB0")
```

Load R source

```{r source, cache = FALSE}
source(here::here("code/color.R"))
```

## Load data

Load site stat each super-pop

```{r load-site-stat-eachPop, message=FALSE, warning=FALSE}
data_dir = '/niuyw-usb-disk/Projects/STR/str_stat'

sample_stat_byPop = read_csv(file.path(data_dir, "sample_stat_byPop.csv"))
```

Load sample stat

```{r load-sam-stat, message=FALSE, warning=FALSE}
sam_stat = readRDS(here::here("output/sample_stat", "sam_stat.rds"))
```

Load site/allele list of each super-pop

```{r load-site-allele-lst-each-sPop, message=FALSE, warning=FALSE}
site_lst = readRDS(here::here("output/pop_sharing", "site_lst.rds"))
lengths(site_lst)
allele_lst = readRDS(here::here("output/pop_sharing", "allele_lst.rds"))
lengths(allele_lst)
```

## pSTR stat of each dataset

Dataset

```{r , message=FALSE, warning=FALSE}
DT::datatable(sample_stat_byPop)
```

Plot

```{r site-stat-each-dataset, fig.width=3, fig.height=2.5}
super_pops = c("AFR", "AMR", "EAS", "EUR", "SAS", "NyuWa")

p.spopStat = sample_stat_byPop %>%
  filter(dataset %in% c("Merged", "1KGP", "NyuWa", super_pops)) %>%
  dplyr::select(dataset,
                `Di-` = period_2, `Tri-` = period_3, `Tet-` = period_4,
                `Pen-` = period_5, `Hex-` = period_6) %>%
  pivot_longer(cols = c("Di-", "Tri-", "Tet-", "Pen-", "Hex-")) %>%
  mutate(name = factor(name, levels = rev(c("Di-", "Tri-", "Tet-", "Pen-", "Hex-")))) %>%
  ggplot(aes(x = reorder(dataset, value, sum), y = value, fill = name)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0),
                     labels = label_comma()) +
  scale_fill_manual(values = period_color_map) +
  labs(x = NULL, y = "# of pSTRs", fill = NULL) +
  coord_flip() +
  theme(legend.position = c(0.75, 0.25))
p.spopStat

ggsave(p.spopStat, filename = here::here("output", DOCNAME, "Fig.spopStat.pdf"),
       height = 2.5, width = 3)
saveRDS(p.spopStat, file = here::here("output", DOCNAME, "Fig.spopStat.rds"))
```

Each pop

```{r site-stat-each-pop, fig.width=7, fig.height=3}
# order bymedian
tmp_df = sam_stat %>%
  filter(pop != "NA") %>%
  group_by(super_pop) %>%
  mutate(m1 = median(tot)) %>%
  group_by(pop) %>%
  mutate(m2 = median(tot)) %>%
  mutate(m2  = case_when(
    pop == "IBS" ~ 32775+1, # since IBS/TSI have the same median value
    TRUE ~ m2
  )) %>%
  arrange(-m1, -m2, -tot) %>%
  dplyr::select(-m1, -m2)

p.popStat = sample_stat_byPop %>%
  filter(!(dataset %in% c("Merged", "1KGP", "NyuWa", super_pops))) %>%
  mutate(dataset = factor(dataset, levels = unique(tmp_df$pop))) %>%
  dplyr::select(dataset,
                `Di-` = period_2, `Tri-` = period_3, `Tet-` = period_4,
                `Pen-` = period_5, `Hex-` = period_6) %>%
  pivot_longer(cols = c("Di-", "Tri-", "Tet-", "Pen-", "Hex-")) %>%
  mutate(name = factor(name, levels = rev(c("Di-", "Tri-", "Tet-", "Pen-", "Hex-")))) %>%
  ggplot(aes(x = dataset, y = value, fill = name)) +
  geom_col() +
  scale_y_continuous(expand = c(0, 0),
                     labels = label_comma()) +
  scale_fill_manual(values = period_color_map) +
  labs(y = NULL, x = "# of pSTRs", fill = NULL)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
p.popStat

ggsave(p.popStat, filename = here::here("output", DOCNAME, "Fig.popStat.pdf"),
       height = 2.5, width = 4)
saveRDS(p.popStat, file = here::here("output", DOCNAME, "Fig.popStat.rds"))
```

## NyuWa and 1KGP

### Site venn

```{r fig.width=4, fig.height=4}
num = list(A = 59411, B = 89719, AB = 216883)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.3, "1KGP", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)

pdf(here::here("output", DOCNAME, "Fig.cmp_NyuWa_1KGP.site.pdf"), height = 4, width = 4)
num = list(A = 59411, B = 89719, AB = 216883)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.3, "1KGP", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)
dev.off()
```

### Allele venn

```{r fig.width=4, fig.height=4}
num = list(A = 262273, B = 523063, AB = 774479)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.3, "1KGP", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)

pdf(here::here("output", DOCNAME, "Fig.cmp_NyuWa_1KGP.allele.pdf"), height = 4, width = 4)
num = list(A = 262273, B = 523063, AB = 774479)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.3, "1KGP", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)
dev.off()
```

## NyuWa and EAS

Number

```{r}
length(intersect(site_lst$NyuWa, site_lst$EAS))
length(intersect(allele_lst$NyuWa, allele_lst$EAS))
```

### Site venn

```{r fig.width=4, fig.height=4}
num = list(A = 9322, B = 175028, AB = 131574)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.2, "EAS", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)

pdf(here::here("output", DOCNAME, "Fig.cmp_NyuWa_1KGP_EAS.site.pdf"), height = 4, width = 4)
num = list(A = 9322, B = 175028, AB = 131574)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.2, "1KGP EAS", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)
dev.off()
```

### Allele venn

```{r fig.width=4, fig.height=4}
num = list(A = 37977, B = 825162, AB = 472380)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.3, "EAS", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)

pdf(here::here("output", DOCNAME, "Fig.cmp_NyuWa_1KGP_EAS.allele.pdf"), height = 4, width = 4)
num = list(A = 37977, B = 825162, AB = 472380)
v = venneuler(c(A = num$A, B = num$B, "A&B" = num$AB))
v$labels <- c("","","")
plot(v)
text(v$centers[1, 1] + 0.2, v$centers[1, 2], prettyNum(num$A, big.mark = ","), cex = 1) #A
text(v$centers[2, 1] - 0.2, v$centers[2, 2], prettyNum(num$B, big.mark = ","), cex = 1) #B
text(0.5, 0.5, prettyNum(num$AB, big.mark = ","), cex = 1) #AB
text(v$centers[1, 1] + 0.3, v$centers[1, 2] + 0.3, "EAS", cex = 1)
text(v$centers[2, 1] - 0.3, v$centers[2, 2] + 0.3, "NyuWa", cex = 1)
dev.off()
```

## AF comparison

Load site list of each super-pop

```{r load-site-lst-each-sPop, message=FALSE, warning=FALSE}
data_dir = '/niuyw-usb-disk/Projects/STR/str_stat'

super_pops = c("1KGP", "AFR", "AMR", "EAS", "EUR", "SAS", "NyuWa")
site_lst = lapply(super_pops, function(x){
  d = read_tsv(file.path(data_dir, "sample_stat_byPop", paste0("site_lst.", x)), col_names = FALSE)
  d$X1
})
names(site_lst) = super_pops
```

Load STR MAF

```{r load-str-maf, message=FALSE, warning=FALSE}
str_maf = read_csv("/T02Data/niuyw/niuyw-usb-disk/Projects/STR/str_stat/str_maf.csv")
```

### NyuWa vs. 1KGP

```{r , fig.height=3.5, fig.width=3.5}
common_loci = intersect(site_lst$`1KGP`, site_lst$NyuWa)
d4p = str_maf %>%
  filter(site %in% common_loci)

p.af_cmp.all = d4p %>%
  ggplot(aes(x = maf_NyuWa, y = maf_1KGP)) +
  geom_point_rast(alpha = .3, raster.dpi = 300, size = .5) + 
  stat_cor(method = "pearson") +
  labs(x = 'MAF (NyuWa)', y = 'MAF (1KGP)') +
  geom_abline(slope = 1, intercept = 0, color = 'blue')
p.af_cmp.all
```

### NyuWa vs. EAS

```{r , fig.height=3.5, fig.width=3.5}
common_loci = intersect(site_lst$EAS, site_lst$NyuWa)
d4p = str_maf %>%
  filter(site %in% common_loci)

p.af_cmp.eas = d4p %>%
  ggplot(aes(x = maf_NyuWa, y = maf_EAS)) +
  geom_point_rast(alpha = .3, raster.dpi = 300, size = .5) + 
  stat_cor(method = "pearson") +
  labs(x = 'MAF (NyuWa)', y = 'MAF (1KGP EAS)') +
  geom_abline(slope = 1, intercept = 0, color = 'blue')
p.af_cmp.eas
```

### NyuWa vs. AFR

```{r , fig.height=3.5, fig.width=3.5}
common_loci = intersect(site_lst$AFR, site_lst$NyuWa)
d4p = str_maf %>%
  filter(site %in% common_loci)

p.af_cmp.afr = d4p %>%
  ggplot(aes(x = maf_NyuWa, y = maf_AFR)) +
  geom_point_rast(alpha = .3, raster.dpi = 300, size = .5) + 
  stat_cor(method = "pearson") +
  labs(x = 'MAF (NyuWa)', y = 'MAF (1KGP AFR)') +
  geom_abline(slope = 1, intercept = 0, color = 'blue')
p.af_cmp.afr
```

### Align plots

```{r fig.width=5, fig.height=2.5}
p.af_cmp = p.af_cmp.all + p.af_cmp.eas +
  plot_layout(ncol = 2)
p.af_cmp

ggsave(p.af_cmp,
  filename = here::here("output", DOCNAME, "Fig.dataset_cmp_AF.pdf"),
  height = 3.5, width = 7
)
saveRDS(p.af_cmp, file = here::here("output", DOCNAME, "Fig.dataset_cmp_AF.rds"))
```

## Allele freq (NyuWa and 1KGP)

Load allele info

```{r load-allele-info, message=FALSE, warning=FALSE}
allele_info.nyuwa <- read_tsv("/niuyw-usb-disk/Projects/STR/str_mutational/allele_info.nyuwa.txt") %>%
  mutate(motif_len = period) %>%
  mutate(period = case_when(
    period == 2 ~ "Di-",
    period == 3 ~ "Tri-",
    period == 4 ~ "Tet-",
    period == 5 ~ "Pen-",
    period == 6 ~ "Hex-"
  )) %>%
  mutate(period = factor(period, levels = c("Di-", "Tri-", "Tet-", "Pen-", "Hex-")))
allele_info.1kgp <- read_tsv("/niuyw-usb-disk/Projects/STR/str_mutational/allele_info.1kgp.txt") %>%
  mutate(motif_len = period) %>%
  mutate(period = case_when(
    period == 2 ~ "Di-",
    period == 3 ~ "Tri-",
    period == 4 ~ "Tet-",
    period == 5 ~ "Pen-",
    period == 6 ~ "Hex-"
  )) %>%
  mutate(period = factor(period, levels = c("Di-", "Tri-", "Tet-", "Pen-", "Hex-")))
```

Plot

```{r fig.width=7, fig.height=2.5}
p.af.nyuwa <- allele_info.nyuwa %>%
  mutate(bin = cut(af,
    c(
      0, 0.001, 0.01,
      0.1, 0.5, 0.9,
      0.99, 0.999, 1
    ),
    labels = c(
      "0.001", "0.01",
      "0.1", "0.5", "0.9",
      "0.99", "0.999", "1"
    )
  )) %>%
  group_by(period, bin) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = bin, y = prop, fill = period)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = cPalette) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Allele frequency bins", y = "% of alleles",
    #title = "NyuWa",
    fill = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = c(.8, .8)
  )

p.af.1kgp <- allele_info.1kgp %>%
  mutate(bin = cut(af,
    c(
      0, 0.001, 0.01,
      0.1, 0.5, 0.9,
      0.99, 0.999, 1
    ),
    labels = c(
      "0.001", "0.01",
      "0.1", "0.5", "0.9",
      "0.99", "0.999", "1"
    )
  )) %>%
  group_by(period, bin) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = bin, y = prop, fill = period)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = cPalette) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Allele frequency bins", y = "% of alleles",
    #title = "1KGP",
    fill = NULL
  ) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = c(.8, .8)
  )

p.af = p.af.nyuwa + p.af.1kgp +
  plot_layout(ncol = 2)
p.af
```

## Call rate

Load

```{r load-callrate, message=FALSE, warning=FALSE}
nyuwa = read_tsv("/niuyw-usb-disk/Projects/STR/cmp_NyuWa_1KGP/specific_nyuwa.txt", col_names = FALSE)
mean(nyuwa$X1*100)

kgp = read_tsv("/niuyw-usb-disk/Projects/STR/cmp_NyuWa_1KGP/specific_1kgp.txt", col_names = FALSE)
mean(kgp$X1*100)
```

Plot

```{r fig.width=7, fig.height=2.5}
p1 = nyuwa %>%
  ggplot(aes(x = X1*100)) +
  geom_histogram(binwidth = 1, fill = "#1f77b4") +
  geom_vline(xintercept = mean(nyuwa$X1*100), linetype = "dashed") +
  scale_y_log10(
    expand = c(0, 0),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  labs(x = "Loci call rate of 1KGP dataset (%)", y = "# of pSTR",
       title = "NyuWa specific pSTRs")
p2 = kgp %>%
  ggplot(aes(x = X1*100)) +
  geom_histogram(binwidth = 1, fill = "#1f77b4") +
  geom_vline(xintercept = mean(kgp$X1*100), linetype = "dashed") +
  scale_y_log10(
    expand = c(0, 0),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  labs(x = "Loci call rate of NyuWa dataset (%)", y = "# of pSTR",
       title = "1KGP specific pSTRs")
p.callrate = p1 + p2 +
  plot_layout(ncol = 2)
p.callrate

ggsave(p.callrate,
  filename = here::here("output", DOCNAME, "Fig.dataset_specific_loci.callrate.pdf"),
  height = 3.5, width = 7
)
saveRDS(p.callrate, file = here::here("output", DOCNAME, "Fig.dataset_specific_loci.callrate.rds"))
```

## Align plots

```{r fig.width=7, fig.height=8.5}
p = p.af/p.callrate/p.af_cmp +
  plot_layout(heights = c(2.5, 2.5, 3.5))
p

ggsave(p,
  filename = here::here("output", DOCNAME, "Fig.dataset_cmp.pdf"),
  height = 8.5, width = 7
)
```

## Session info

