---
title: "FC.1KGP_EAS.pops.het01"
---

```{r knitr, include = FALSE}
DOCNAME <- "FC.1KGP_EAS.pops.het01"
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

NOW <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    NOW <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - NOW)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache.lazy     = FALSE,
  cache.comments = FALSE,
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 10,
  fig.height     = 8,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```

## Setup

Load packages

```{r libaries, cache = FALSE, message=FALSE, warning=FALSE}
# Tidyverse
library(tidyverse)

# plot
library(scales)
library(patchwork)
library(cowplot)
library(ggpubr)
library(ggcorrplot)
library(ggrepel)
library(ggrastr)
library(EnvStats)
theme_set(theme_cowplot(
  font_size = 10, # axis.title
  rel_small = 8 / 10, # axis.text
  rel_tiny = 9 / 10,
  rel_large = 10 / 10
))
geom.text.size <- 8 / (14 / 5)

# heatmap
library(pheatmap)

# color
library(ggsci)
library(RColorBrewer)

# data.table
library(data.table)
```

Load R source

```{r source, cache = FALSE}
source(here::here("code/color.R"))
```

Load rds (fast)

```{r load-rds, cache=FALSE}
dat_lst <- readRDS(here::here("output", DOCNAME, "FC.rds"))
```

## Load and neat

Load STR info

```{r message=FALSE, warning=FALSE, eval=FALSE}
str_info <- read_tsv("/T02Data/niuyw/niuyw-usb-disk/Projects/STR/str_info/str_info.txt") %>%
  filter(het > 0.1) %>%
  mutate(csq2 = case_when(
    csq == "upstream" ~ "Promoter",
    csq == "nc_intron" ~ "Noncoding_intron",
    csq == "intron" ~ "Coding_intron",
    csq == "downstream" ~ "Downstream",
    csq == "UTR3" ~ "3'UTR",
    csq == "intergenic" ~ "Intergenic",
    csq == "nc_exon" ~ "Noncoding_exon",
    csq == "UTR5" ~ "5'UTR",
    TRUE ~ csq
  ))
dim(str_info)
```

Load FC and p values.

```{r load-dat, message=FALSE, warning=FALSE, eval=FALSE}
data_dir <- "/niuyw-usb-disk/Projects/STR/fold_change"
chroms <- paste0("chr", 1:22)

# pops to cmp
cmp_pops <- c(
  "CHB.1KGPvsCHS.1KGP", "CDXvsCHS.1KGP", "KHVvsCHS.1KGP", "JPTvsCHS.1KGP",
  "CHB.1KGPvsJPT", "CDXvsJPT", "KHVvsJPT",
  "CHB.1KGPvsKHV", "CDXvsKHV",
  "CHB.1KGPvsCDX"
)

# load
dat_lst <- lapply(cmp_pops, function(x) {
  tmp_lst <- lapply(chroms, function(y) {
    read_tsv(file.path(data_dir, x, paste0(y, ".FC.txt"))) %>%
      mutate(site = str_match(site, ".*?:.*?:")) %>%
      mutate(site = str_remove(site, ":$")) %>%
      filter(site %in% str_info$site)
  })
  df <- do.call(rbind, tmp_lst)
  # get q-value
  df <- df %>%
    mutate(q = p.adjust(ranksums.p, method = "BH")) %>%
    left_join(str_info, by = "site") %>%
    arrange(q, -abs(LFC))
  df
})
names(dat_lst) <- cmp_pops
```

## Stats

0.01

```{r}
lapply(dat_lst, FUN = function(x) {
  x %>%
    filter(q < 0.01) %>%
    mutate(change = case_when(
      LFC > 0 ~ "Up",
      TRUE ~ "Dn"
    )) %>%
    pull(change) %>%
    table()
})
```

0.05

```{r}
lapply(dat_lst, FUN = function(x) {
  x %>%
    filter(q < 0.05) %>%
    mutate(change = case_when(
      LFC > 0 ~ "Up",
      TRUE ~ "Dn"
    )) %>%
    pull(change) %>%
    table()
})
```

## Show {.tabset}

Top 100

```{r show-top100, results = "hide"}
# pops to cmp
cmp_pops <- names(dat_lst)
src_list <- lapply(cmp_pops, function(s) {
  src <- c(
    "### {{s}} {.unnumbered}",
    "```{r show-top100-{{s}}}",
    "df = dat_lst[['{{s}}']]",
    "df_q005 = df %>% dplyr::filter(q < .05)",
    "dim(df_q005)",
    "DT::datatable(head(df_q005, 100), filter = 'top')",
    "```",
    ""
  )
  # print(src)
  knitr::knit_expand(text = src)
})

out <- knitr::knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

## Plot

Plot

```{r fig.width=8, fig.height=8}
# pops to cmp
cmp_pops <- names(dat_lst)

plot_volcano <- function(dat = NULL) {
  dat %>%
    ggplot(aes(x = LFC, y = -log10(q))) +
    geom_point_rast(size = 0.1, raster.dpi = 300, color = "grey50") +
    geom_hline(
      yintercept = -log10(0.01),
      linetype = "dashed",
      color = "gray"
    ) +
    xlim(-1, 1) +
    labs(
      x = expression(Log[2] ~ fold ~ change),
      y = expression(-Log[10] ~ p - value)
    )
}

# plot
p_lst <- lapply(cmp_pops, function(x) {
  # print(x)
  p <- plot_volcano(dat_lst[[x]])
  p
})
# p_lst[[1]]

# layout
layout <- c(
  area(t = 1, l = 1),
  area(t = 1, l = 2),
  area(t = 1, l = 3),
  area(t = 1, l = 4),
  area(t = 2, l = 1),
  area(t = 2, l = 2),
  area(t = 2, l = 3),
  area(t = 3, l = 1),
  area(t = 3, l = 2),
  area(t = 4, l = 1)
)
# plot(layout)

p <- wrap_plots(p_lst) +
  plot_layout(design = layout)
p

ggsave2(p,
  filename = here::here("output", DOCNAME, "Fig.FC_1KGP_EAS_pops.pdf"),
  height = 8, width = 8
)
saveRDS(p, file = here::here("output", DOCNAME, "Fig.FC_1KGP_EAS_pops.rds"))
```

## Output

```{r write-tsv, eval=FALSE}
xxx <- lapply(cmp_pops, function(x) {
  dat_lst[[x]] %>%
    write_tsv(file = here::here("output", DOCNAME, paste0(x, ".FC.tsv")))
})

xxx <- lapply(cmp_pops, function(x) {
  dat_lst[[x]] %>%
    filter(q < 0.05) %>%
    write_tsv(file = here::here("output", DOCNAME, paste0(x, ".FC_q005.tsv")))
})
```

## Save

```{r save, cache=FALSE, eval=FALSE}
saveRDS(dat_lst, file = here::here("output", DOCNAME, "FC.rds"))
```

## Session info

